---
title: "Generalized IPW"
author: "Xiaoting Chen"
date: "2024-05-24"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is translated from the Generalized_IPW.ipynb file and expanded on the spline function and prediction curve generation.\


## Prep

### packages
```{r, warning=FALSE, message=FALSE}
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(stats) # handle normal distribution calculations
library(splines) # using splines
library(mgcv)
library(DescTools) # calculate mode
library(tableone)
library(cobalt)
```

### data

```{r}
all_combos <- readRDS("all_combos.rds")
# in generalized IPW estimation, M is not considered, thus we only keep all the Y&X combos
all_combos <- all_combos[seq(1, length(all_combos), by = 5)]
```


```{r}
mesa_std <- read.csv('../../data_processed/MESA/mesa_std.csv')

analysis_feature <- c('cvd_10y_HF', 'cvd_10y_noHF', 'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp',
                      'site', 'race')

mesa_std <- mesa_std %>%
  select(analysis_feature) %>%
  mutate( 
    Diabetes = as.factor(Diabetes),
         site = as.factor(site),
         race = as.factor(race)) 

```

```{r}
jhs_std <- read.csv('../../data_processed/JHS/jhs_std.csv')

analysis_feature <- c('cvd_10y_HF', 'cvd_10y_noHF', 'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp')

jhs_std <- jhs_std %>%
  select(analysis_feature) %>%
  mutate(
    Diabetes = as.factor(Diabetes)) 
```

```{r}
named_datasets <- list(mesa = mesa_std, jhs = jhs_std)
names(named_datasets) <- c("mesa", "jhs")
```

### function    
```{r}
conditional_densities <- function(data, treatment, formula_ps_no_con, formula_ps_con, use_confounders = TRUE) {

  formula <- if(use_confounders) formula_ps_con else formula_ps_no_con
  
  # Fit the linear model
  model <- lm(as.formula(formula), data = data)
  
  # Calculate the fitted values and standard deviation of residuals
  fitted_values <- fitted(model)
  resid_std <- sd(resid(model))
  
  # Calculate the density of treatment under a normal distribution with parameters from the model
  densities <- dnorm(data[[treatment]], mean = fitted_values, sd = resid_std)
  
  # Return the densities as a vector indexed similarly to the fitted values
  return(setNames(densities, names(fitted_values)))
}

```


## estimated using generalized IPW

### original sample
```{r, warning=FALSE}
report_df <- data.frame(
  Data = character(),
  Y = numeric(),
  X = numeric(),
  intercept = numeric(),
  intercept_pvalue = numeric(),
  X_coef = numeric(),
  X_pvalue = numeric(),
  X_ci_lwr = numeric(),
  X_ci_upr = numeric(),
  stringsAsFactors = FALSE 
)


for (combo in all_combos) {
  
  # get values
  data_name = combo$data
  data <- named_datasets[[data_name]]
  
  Y <- combo$Y
  X <- combo$X
  Z <- combo$covariates
  
  # generate formula
  formula_ps_no_str <- paste(X, "~ 1")
  formula_ps_no <- as.formula(formula_ps_no_str)
    
  formula_ps_str <- paste(X, "~ ", paste(Z, collapse=" + "))
  formula_ps <- as.formula(formula_ps_str)
    
  formula_outcome_str <- paste(Y, "~",X)
  formula_outcome <- as.formula(formula_outcome_str)
  
  # estimate
  denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
  numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
  propensity_density = numerator / denominator
  threshold <- quantile(propensity_density, 0.99)
  data <- data[propensity_density <= threshold, ]
  propensity_density <- propensity_density[propensity_density <= threshold]
  outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
  summary_model <- summary(outcome_mod)

  # store model result
  coefficients <- summary_model$coefficients
  ci <- confint(outcome_mod)   ## CI from the SE of coefs
  
  intercept <- round(coefficients["(Intercept)", "Estimate"], digits = 5)  ## common parts
  intercept_pvalue <- round(coefficients["(Intercept)", "Pr(>|z|)"], digits = 5)
  
  X_coef <- round(coefficients[X, "Estimate"], digits = 5)
  X_pvalue <- round(coefficients[X, "Pr(>|z|)"], digits = 5)
  X_ci_lwr <- round(ci[X,"2.5 %"],digits = 5)
  X_ci_upr <- round(ci[X,"97.5 %"],digits = 5)
  
  new_df <- data.frame(
              Data = data_name,
              Y = Y,
              X = X,
              intercept = intercept,
              intercept_pvalue = intercept_pvalue,
              X_coef = X_coef,
              X_pvalue = X_pvalue,
              X_ci_lwr = X_ci_lwr,
              X_ci_upr = X_ci_upr,
              stringsAsFactors = FALSE )

  report_df <- rbind(report_df, new_df)

}

```

```{r}
write.csv(report_df, "../../results/reordered_results/generalized_ipw/report_df_origin.csv", row.names = FALSE)
```


### bootstrap 
```{r, warning=FALSE, message=FALSE}
n_boot <- 1000
set.seed(1)

report_df_boot <- data.frame(
  boot = numeric(),
  Data = character(),
  Y = numeric(),
  X = numeric(),
  intercept = numeric(),
  intercept_pvalue = numeric(),
  X_coef = numeric(),
  stringsAsFactors = FALSE 
)


for (i in seq_along(all_combos)) {
  
  print(i)  
  combo <- all_combos[[i]]
  
  # get values
  data_name = combo$data
  data <- named_datasets[[data_name]]
  
  Y <- combo$Y
  X <- combo$X
  Z <- combo$covariates
  
  # generate formula
  formula_ps_no_str <- paste(X, "~ 1")
  formula_ps_no <- as.formula(formula_ps_no_str)
    
  formula_ps_str <- paste(X, "~ ", paste(Z, collapse=" + "))
  formula_ps <- as.formula(formula_ps_str)
    
  formula_outcome_str <- paste(Y, "~", X)
  formula_outcome <- as.formula(formula_outcome_str)
 
  for (j in 1:n_boot) {
    # Resample 
    indices <- sample(1:nrow(data), replace = TRUE)
    boot_data <- data[indices, ]
    
    # estimate
    denominator = conditional_densities(boot_data, X, formula_ps_no,formula_ps, use_confounders=T)
    numerator = conditional_densities(boot_data,X, formula_ps_no,formula_ps, use_confounders=F)
    propensity_density = numerator / denominator
    threshold <- quantile(propensity_density, 0.99)
    boot_data <- boot_data[propensity_density <= threshold, ]
    propensity_density <- propensity_density[propensity_density <= threshold]
    outcome_mod <- glm(formula = formula_outcome, data = boot_data, family = binomial(), weights = propensity_density)
    summary_model <- summary(outcome_mod)
   
    # store model result
    coefficients <- summary_model$coefficients
    
    intercept <- round(coefficients["(Intercept)", "Estimate"], digits = 5)  ## common parts
    intercept_pvalue <- round(coefficients["(Intercept)", "Pr(>|z|)"], digits = 5)
    
    X_coef <- round(coefficients[X, "Estimate"], digits = 5)
    
    new_df <- data.frame(
                boot = j,
                Data = data_name,
                Y = Y,
                X = X,
                intercept = intercept,
                intercept_pvalue = intercept_pvalue,
                X_coef = X_coef,
                stringsAsFactors = FALSE )
  
    report_df_boot <- rbind(report_df_boot, new_df)
  }

}

```

```{r}
write.csv(report_df_boot, "../../results/reordered_results/generalized_ipw/report_df_boot.csv", row.names = FALSE)
```


handle extreme values in mesa nFavFood coef
```{r}
# check values
nFavFood_coef <- report_df_boot[(report_df_boot$Data=="mesa")&(report_df_boot$X=="nFavFood") , "X_coef"]
# View(data.frame(nFavFood_coef)) 
# 14 extreme values

# replace with median value
report_df_modified <- report_df_boot
median_value <- median(report_df_modified[(report_df_modified$Data == "mesa") & 
                                      (report_df_modified$X == "nFavFood"), "X_coef"], na.rm = TRUE)
report_df_modified[(report_df_modified$Data == "mesa") & 
               (report_df_modified$X == "nFavFood") & 
               (abs(report_df_modified$X_coef) > 100), "X_coef"] <- median_value
```


summarize from the bootstrap data
```{r}
# summarize
boot_summarized <- report_df_modified %>%
  group_by(Data, Y, X) %>%
  summarise(intercept = mean(intercept),
            X_coef_mean = mean(X_coef),
            X_ci_lwr = quantile(X_coef, 0.025),
            X_ci_upr = quantile(X_coef, 0.975)) %>%
  mutate(
    X_coef = sprintf("%.3f [%.3f, %.3f]", X_coef_mean, X_ci_lwr, X_ci_upr),
    )
```

```{r}
write.csv(boot_summarized, "../../results/reordered_results/generalized_ipw/boot_summarized.csv", row.names = FALSE)
```


## plot prediction curve for significant results
```{r}
boot_summarized <- read.csv("../../results/reordered_results/generalized_ipw/boot_summarized.csv")

sig_from_boot <- boot_summarized %>%
  filter((Y == "cvd_10y_HF") & (X_ci_lwr * X_ci_upr > 0))
# jhs; cvd_10y_HF; nRS
# jhs; cvd_10y_HF; nSES
# mesa; cvd_10y_HF; nSES

report_df_boot <- read.csv("../../results/reordered_results/generalized_ipw/report_df_boot.csv") %>%
  mutate(Data = as.character(Data), Y = as.character(Y), X = as.character(X))
```


### plot significant effects by bootstrap coefs
```{r}
for (i in 1:nrow(sig_from_boot)) {
  print(i)

  # extract significant effects
  data_name <- as.character(sig_from_boot[i, "Data"])
  data <- named_datasets[[data_name]]
  Y_name <- as.character(sig_from_boot[i, "Y"])
  X_name <- as.character(sig_from_boot[i, "X"])
  print(paste("Filtering for Data:", data_name, "Y:", Y_name, "X:", X_name))
  
  # extract bootstrap values
  boot_data <- report_df_boot %>%
    filter((Data == data_name) & (Y == Y_name) & (X == X_name))
  boot_intercepts <- boot_data$intercept
  boot_X_coef <- boot_data$X_coef
  
  print(mean(boot_X_coef))
  
  # prediction 
  ## using the mean intercept and X_coef from bootstrap
  ## and +- 2sd of X_mean as range of predictor values
  x_vals <- seq(from = mean(data[[X_name]])-2*sd(data[[X_name]]),
                to = mean(data[[X_name]])+2*sd(data[[X_name]]), by = 0.1)
  
  pred_probs_bootstrap <- mapply(
    function(intercept, coef) {
      logit <- intercept + coef * x_vals
      1 / (1 + exp(-logit))}, 
    boot_intercepts, 
    boot_X_coef)

  # Calculate mean predicted probabilities and 95% CI
  mean_pred_probs <- apply(pred_probs_bootstrap, 1, mean)
  lower_pred_probs <- apply(pred_probs_bootstrap, 1, quantile, probs = 0.025)
  upper_pred_probs <- apply(pred_probs_bootstrap, 1, quantile, probs = 0.975)
  
  # Combine the data into a data frame for plotting
  pred_data <- data.frame(
    X = x_vals,
    fit = mean_pred_probs,
    lwr = lower_pred_probs,
    upr = upper_pred_probs)

  # plot
  plt <- ggplot(pred_data, aes(x = X)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.2) +  # Confidence interval
    geom_line(aes(y = fit), color = "blue") +  # Mean predicted probabilities
    geom_hline(yintercept = 0, color = "black") +  # Horizontal line at y = 0
    labs(x = paste("Neighborhood", 
                   ifelse(X_name == "nSES", "Socioeconomic Status", "Racial Segregation"),
                   "Level in", data_name),
         y = "Predicted Probabilities of Experiencing CVD") +  # Axis labels
    theme_minimal() +
    theme(
        axis.text.x = element_text(size = 14),  # Increase X-axis tick label size
        axis.text.y = element_text(size = 14),  # Increase Y-axis tick label size
        axis.title.x = element_text(size = 16),  # Increase X-axis label size
        axis.title.y = element_text(size = 16)   # Increase Y-axis label size
    )

  # export plot
  ggsave(paste0(data_name,"_",Y_name,"_",X_name,".png"), plot = plt,
           path = "../../results/reordered_results/generalized_ipw",
           width = 8, height = 6, dpi = 300)
}

```


## Check assumption

dichotomize the neighborhood exposure variables after weight calculation, to examine overlap and balance\\
tutorial from https://ehsanx.github.io/psw/s3.html\\

rename variables for plotting
```{r}
names(mesa_std) <- names(mesa_std) %>%
  dplyr::recode("Diabetes" = "diabetes",
                "hdl" = "HDL",
                "totchol" = "Total_Cholesterol",
                "sbp" = "SBP")

names(jhs_std) <- names(jhs_std) %>%
  dplyr::recode("Diabetes" = "diabetes",
                "hdl" = "HDL",
                "totchol" = "Total_Cholesterol",
                "sbp" = "SBP")
```

MESA nSES
```{r, warning=FALSE}
data <- mesa_std
Y <- "cvd_10y_HF"
X <- "nSES"
Z <- c("age","diabetes","HDL","Total_Cholesterol","SBP","gender",
       "race","site")

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ age+diabetes+HDL+Total_Cholesterol+SBP+gender+race+site")
  
# get weights
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density


smd_after <- bal.tab(
  data[, Z], 
  treat = "nSES", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
                 threshold = 0.1) +
  theme(axis.text = element_text(size = 15),  
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 15),  
        legend.title = element_text(size = 15)) +  
  scale_x_continuous(name = "Treatment-Covariate Correlations for n-SES in MESA") +  
  theme(plot.title = element_blank(),  
        plot.subtitle = element_blank())  

plt

ggsave("mesa_nses_balance.png", plot = plt,
           path = "../../results/reordered_results/generalized_ipw",
           width = 8, height = 6, dpi = 300)
```


JHS nSES
```{r}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nSES"
Z <- c("age","diabetes","HDL","Total_Cholesterol","SBP","gender")

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ age+gender+diabetes+HDL+Total_Cholesterol+SBP")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density

smd_after <- bal.tab(
  data[, Z], 
  treat = "nSES", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
                 threshold = 0.1) +
  theme(axis.text = element_text(size = 15),  
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 15),  
        legend.title = element_text(size = 15)) +  
  scale_x_continuous(name = "Treatment-Covariate Correlations for n-SES in JHS") +  
  theme(plot.title = element_blank(),  
        plot.subtitle = element_blank())   

plt
ggsave("jhs_nses_balance.png", plot = plt,
           path = "../../results/reordered_results/generalized_ipw",
           width = 8, height = 6, dpi = 300)
```


JHS nRS
```{r}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nRS"
Z <- c("age","diabetes","HDL","Total_Cholesterol","SBP","gender")

# generate formula 
formula_ps_no <- as.formula("nRS ~ 1")
formula_ps <- as.formula("nRS ~ age+gender+diabetes+HDL+Total_Cholesterol+SBP")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density

# using continuous nSES
smd_after <- bal.tab(
  data[, Z], 
  treat = "nRS", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
                 threshold = 0.1) +
  theme(axis.text = element_text(size = 15),  
        axis.title = element_text(size = 18),
        legend.text = element_text(size = 15),  
        legend.title = element_text(size = 15)) +  
  scale_x_continuous(name = "Treatment-Covariate Correlations for n-Racial Segregation in JHS") +  
  theme(plot.title = element_blank(),  
        plot.subtitle = element_blank())  

plt

ggsave("jhs_nrs_balance.png", plot = plt,
           path = "../../results/reordered_results/generalized_ipw",
           width = 8, height = 6, dpi = 300)
```




