---
title: "Mediation analysis"
author: "Xiaoting Chen"
date: "2024-08-22"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Prep

### packages
```{r, warning=FALSE}
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(stats) # handle normal distribution calculations
library(splines) # using splines
library(mgcv)
library(bda) # sobel test
```

### data

```{r}
all_combos <- readRDS("all_combos.rds")
```

```{r}
analysis_feature_mesa <- c('cvd_10y_HF', 'cvd_10y_noHF', 'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp',
                      'site', 'race')

analysis_feature_jhs <- c('cvd_10y_HF', 'cvd_10y_noHF', 
                      'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp')

ind_feature_map <- c('0' = 1, '1' = 2, '2' = 3)
```

```{r}
mesa_std <- read.csv('../../data_processed/MESA/mesa_std.csv')

mesa_std <- mesa_std %>%
  select(analysis_feature_mesa) %>%
  mutate(# # treat ind behavirol variables as conitnuous, avoid partial significant 
    Diabetes = as.factor(Diabetes),
         site = as.factor(site),
         race = as.factor(race)) %>%
  mutate( # recode to avoid 0
    nutrition = recode(nutrition, !!!ind_feature_map),
    PhysAct = recode(PhysAct, !!!ind_feature_map),
    currentSmoker = recode(currentSmoker, !!!ind_feature_map),
    alc = recode(alc, !!!ind_feature_map)
  )

```

```{r}
summary(mesa_std)
```

```{r}
jhs_std <- read.csv('../../data_processed/JHS/jhs_std.csv')

jhs_std <- jhs_std %>%
  select(analysis_feature_jhs) %>%
  mutate(Diabetes = as.factor(Diabetes)) %>%
  mutate( # recode to avoid 0
    nutrition = recode(nutrition, !!!ind_feature_map),
    PhysAct = recode(PhysAct, !!!ind_feature_map),
    currentSmoker = recode(currentSmoker, !!!ind_feature_map),
    alc = recode(alc, !!!ind_feature_map)
  )
```

```{r}
summary(jhs_std)
```


```{r}
named_datasets <- list(mesa = mesa_std, jhs = jhs_std)
names(named_datasets) <- c("mesa", "jhs")
```



## result generation

**Steps:**
logitY ~ b0 + b1X + bZ + e
M ~ b0 + b2X + bZ+ e
logitY ~ b0 + b3X + b4M + bZ + e

### bootstrap to obtain ci
```{r, warning=FALSE, message=FALSE}
n_boot <- 1000
set.seed(123)

med_report_df_boot <- data.frame(
  boot = numeric(),
  Data = character(),
  Y = character(),
  X = character(),
  Me = character(),
  b1 = numeric(),
  b2 = numeric(),
  b3 = numeric(),
  b4 = numeric(),
  b2_b4 = numeric(),
  stringsAsFactors = FALSE 
)

for (i in seq_along(all_combos)) {
  
  print(i)  
  combo <- all_combos[[i]]
  
  # get values
  data_name = combo$data
  data <- named_datasets[[data_name]]
  Y_name <- combo$Y
  X_name <- combo$X
  M_name <- combo$M
  Z <- combo$covariates
  
  # generate formula for mediation test
  formula_X_to_Y <- paste(Y_name, "~", X_name, "+", paste(Z, collapse=" + "))
  formula_X_to_Y <- as.formula(formula_X_to_Y)
  formula_X_to_M <- paste(M_name, "~", X_name, "+", paste(Z, collapse=" + "))
  formula_X_to_M <- as.formula(formula_X_to_M)
  formula_XM_to_Y <- paste(Y_name, "~", X_name, "+", M_name, "+", paste(Z, collapse=" + "))
  formula_XM_to_Y <- as.formula(formula_XM_to_Y)

  # Bootstrap to get CIs

  for (j in 1:n_boot) {
    # Resample 
    indices <- sample(1:nrow(data), replace = TRUE)
    boot_data <- data[indices, ]
    
    # fit the models on the bootstrap sample
    boot_mod_X_to_Y <- glm(formula = formula_X_to_Y, data = boot_data, family = binomial())
    boot_mod_X_to_M <- glm(formula = formula_X_to_M, data = boot_data)
    boot_mod_XM_to_Y <- glm(formula = formula_XM_to_Y, data = boot_data, family = binomial())
    
    # Extract coefficients
    boot_coef_X_to_Y <- coef(boot_mod_X_to_Y)
    boot_coef_X_to_M <- coef(boot_mod_X_to_M)
    boot_coef_XM_to_Y <- coef(boot_mod_XM_to_Y)
    
    new_df <- data.frame(
                boot = j,
                Data = data_name,
                Y = Y_name,
                X = X_name,
                Me = M_name,
                b1 = boot_coef_X_to_Y[X_name],
                b2 = boot_coef_X_to_M[X_name],
                b3 = boot_coef_XM_to_Y[X_name],
                b4 = boot_coef_XM_to_Y[M_name], 
                b2_b4 = boot_coef_X_to_M[X_name]*boot_coef_XM_to_Y[M_name],
                stringsAsFactors = FALSE )
  
    med_report_df_boot <- rbind(med_report_df_boot, new_df)
  }
  
}
```

```{r}
write.csv(med_report_df_boot, "../../results/reordered_results/mediation/med_report_df_boot.csv", row.names = FALSE)
```


### summarize mediation result 

```{r}
boot_summarized <- med_report_df_boot %>%
  group_by(Data, Y, X, Me) %>%
  summarise(b1_mean = mean(b1),
            b1_ci_lwr = quantile(b1, 0.025),
            b1_ci_upr = quantile(b1, 0.975),
            b2_mean = mean(b2),
            b2_ci_lwr = quantile(b2, 0.025),
            b2_ci_upr = quantile(b2, 0.975),
            b3_mean = mean(b3),
            b3_ci_lwr = quantile(b3, 0.025),
            b3_ci_upr = quantile(b3, 0.975),
            b4_mean = mean(b4),
            b4_ci_lwr = quantile(b4, 0.025),
            b4_ci_upr = quantile(b4, 0.975),
            b2_b4_mean = mean(b2_b4),
            b2_b4_ci_lwr = quantile(b2_b4, 0.025),
            b2_b4_ci_upr = quantile(b2_b4, 0.975))

write.csv(boot_summarized, "../../results/reordered_results/mediation/med_boot_summarized.csv", row.names = FALSE)
```

```{r}
med_boot_sig <- boot_summarized %>%
  filter(b2_b4_ci_lwr*b2_b4_ci_upr > 0) %>%
  mutate(
    TotalEffect_b1 = sprintf("%.3f [%.3f, %.3f]", b1_mean, b1_ci_lwr, b1_ci_upr),
    b2 = sprintf("%.3f [%.3f, %.3f]", b2_mean, b2_ci_lwr, b2_ci_upr),
    DirectEffect_b3 = sprintf("%.3f [%.3f, %.3f]", b3_mean, b3_ci_lwr, b3_ci_upr),
    b4 = sprintf("%.3f [%.3f, %.3f]", b4_mean, b4_ci_lwr, b4_ci_upr),
    IndirectEffect_b2_b4 = sprintf("%.3f [%.3f, %.3f]", b2_b4_mean, b2_b4_ci_lwr, b2_b4_ci_upr)
  )

write.csv(med_boot_sig, "../../results/reordered_results/mediation/med_boot_sig.csv", row.names = FALSE)
```


### multiple mediator models

combine the multiple mediators identified in med_boot_sig

1. logitY ~ b0 + b1X + bZ + e
2. Me_1 ~ b0 + b21X + bZ+ e
  Me_2 ~ b0 + b22X + bZ+ e
  Me_3 ~ b0 + b23X + bZ+ e
3. logitY ~ b0 + b3X + b41Me_1 + b42Me_2 + b43Me_3 + bZ + e


#### JHS nFavFood -> FamIncome, currentSmoker
```{r, warning=FALSE, message=FALSE}
jhs_nFavFood_multi <- data.frame(
  boot = numeric(),
  b1 = numeric(),
  b21 = numeric(),
  b22 = numeric(),
  b3 = numeric(),
  b41 = numeric(),
  b42 = numeric(),
  stringsAsFactors = FALSE 
)

# model setting
data <- jhs_std
Y_name <- "cvd_10y_HF"
X_name <- "nFavFood"
Me_1 <- "FamIncome"
Me_2 <- "currentSmoker"
Z <- c("age","gender","Diabetes", "hdl","totchol","sbp")
  
# generate formula for mediation test
formula_X_to_Y <- paste(Y_name, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Y <- as.formula(formula_X_to_Y)

formula_X_to_Me_1 <- paste(Me_1, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_1 <- as.formula(formula_X_to_Me_1)

formula_X_to_Me_2 <- paste(Me_2, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_2 <- as.formula(formula_X_to_Me_2)

formula_XM_to_Y <- paste(Y_name, "~", X_name, "+", Me_1, "+", Me_2, "+", paste(Z, collapse=" + "))
formula_XM_to_Y <- as.formula(formula_XM_to_Y)

# Bootstrap to get CIs
n_boot <- 1000
set.seed(123)

for (j in 1:n_boot) {
    # Resample 
    indices <- sample(1:nrow(data), replace = TRUE)
    boot_data <- data[indices, ]
    
    # fit the models on the bootstrap sample
    boot_mod_X_to_Y <- glm(formula = formula_X_to_Y, data = boot_data, family = binomial())
    boot_mod_X_to_Me_1 <- glm(formula = formula_X_to_Me_1, data = boot_data)
    boot_mod_X_to_Me_2 <- glm(formula = formula_X_to_Me_2, data = boot_data)
    boot_mod_XM_to_Y <- glm(formula = formula_XM_to_Y, data = boot_data, family = binomial())
    
    # Extract coefficients
    boot_coef_X_to_Y <- coef(boot_mod_X_to_Y)
    boot_coef_X_to_Me_1 <- coef(boot_mod_X_to_Me_1)
    boot_coef_X_to_Me_2 <- coef(boot_mod_X_to_Me_2)
    boot_coef_XM_to_Y <- coef(boot_mod_XM_to_Y)
    
    new_df <- data.frame(
                boot = j,
                b1 = boot_coef_X_to_Y[X_name],
                b21 = boot_coef_X_to_Me_1[X_name],
                b22 = boot_coef_X_to_Me_2[X_name],
                b3 = boot_coef_XM_to_Y[X_name],
                b41 = boot_coef_XM_to_Y[Me_1],
                b42 = boot_coef_XM_to_Y[Me_2],
                stringsAsFactors = FALSE )
  
    jhs_nFavFood_multi <- rbind(jhs_nFavFood_multi, new_df)
}

write.csv(jhs_nFavFood_multi, "../../results/reordered_results/mediation/multi_mediator/jhs_nFavFood_multi.csv", row.names = FALSE)
```


```{r}
jhs_nFavFood_multi_sum <- jhs_nFavFood_multi %>%
  mutate(b_21_41 = b21*b41,
         b_22_42 = b22*b42) %>%
  summarise(b1_mean = mean(b1),
           b1_ci_lwr = quantile(b1, 0.025),
           b1_ci_upr = quantile(b1, 0.975),
            
            b21_mean = mean(b21),
            b21_ci_lwr = quantile(b21, 0.025),
            b21_ci_upr = quantile(b21, 0.975),
            
            b22_mean = mean(b22),
            b22_ci_lwr = quantile(b22, 0.025),
            b22_ci_upr = quantile(b22, 0.975),
            
            b3_mean = mean(b3),
            b3_ci_lwr = quantile(b3, 0.025),
            b3_ci_upr = quantile(b3, 0.975),
            
            b41_mean = mean(b41),
            b41_ci_lwr = quantile(b41, 0.025),
            b41_ci_upr = quantile(b41, 0.975),
           
           b42_mean = mean(b42),
            b42_ci_lwr = quantile(b42, 0.025),
            b42_ci_upr = quantile(b42, 0.975),
         
            b_21_41_mean = mean(b_21_41),
            b_21_41_ci_lwr = quantile(b_21_41, 0.025),
            b_21_41_ci_upr = quantile(b_21_41, 0.975),
           
           b_22_42_mean = mean(b_22_42),
            b_22_42_ci_lwr = quantile(b_22_42, 0.025),
            b_22_42_ci_upr = quantile(b_22_42, 0.975)) %>%
  mutate(
    TotalEffect_b1 = sprintf("%.3f [%.3f, %.3f]", b1_mean, b1_ci_lwr, b1_ci_upr),
    b21 = sprintf("%.3f [%.3f, %.3f]", b21_mean, b21_ci_lwr, b21_ci_upr),
    b22 = sprintf("%.3f [%.3f, %.3f]", b22_mean, b22_ci_lwr, b22_ci_upr),
    DirectEffect_b3 = sprintf("%.3f [%.3f, %.3f]", b3_mean, b3_ci_lwr, b3_ci_upr),
    b41 = sprintf("%.3f [%.3f, %.3f]", b41_mean, b41_ci_lwr, b41_ci_upr),
    b42 = sprintf("%.3f [%.3f, %.3f]", b42_mean, b42_ci_lwr, b42_ci_upr),
    IndirectEffect_b21_b41 = sprintf("%.3f [%.3f, %.3f]", b_21_41_mean, b_21_41_ci_lwr, b_21_41_ci_upr),
    IndirectEffect_b22_b42 = sprintf("%.3f [%.3f, %.3f]", b_22_42_mean, b_22_42_ci_lwr, b_22_42_ci_upr)
  )

write.csv(jhs_nFavFood_multi_sum, "../../results/reordered_results/mediation/multi_mediator/jhs_nFavFood_multi_sum.csv", row.names = FALSE) 
```

#### JHS nRS -> FamIncome, PhysAct, currentSmoker

```{r, warning=FALSE, message=FALSE}
jhs_nRS_multi <- data.frame(
  boot = numeric(),
  b1 = numeric(),
  b21 = numeric(),
  b22 = numeric(),
  b23 = numeric(),
  b3 = numeric(),
  b41 = numeric(),
  b42 = numeric(),
  b43 = numeric(),
  stringsAsFactors = FALSE 
)

# model setting
data <- jhs_std
Y_name <- "cvd_10y_HF"
X_name <- "nRS"
Me_1 <- "FamIncome"
Me_2 <- "PhysAct"
Me_3 <- "currentSmoker"
Z <- c("age","gender","Diabetes", "hdl","totchol","sbp")
  
# generate formula for mediation test
formula_X_to_Y <- paste(Y_name, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Y <- as.formula(formula_X_to_Y)

formula_X_to_Me_1 <- paste(Me_1, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_1 <- as.formula(formula_X_to_Me_1)

formula_X_to_Me_2 <- paste(Me_2, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_2 <- as.formula(formula_X_to_Me_2)

formula_X_to_Me_3 <- paste(Me_3, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_3 <- as.formula(formula_X_to_Me_3)

formula_XM_to_Y <- paste(Y_name, "~", X_name, "+", Me_1, "+", Me_2, "+", Me_3, "+", paste(Z, collapse=" + "))
formula_XM_to_Y <- as.formula(formula_XM_to_Y)

# Bootstrap to get CIs
n_boot <- 1000
set.seed(123)

for (j in 1:n_boot) {
    # Resample 
    indices <- sample(1:nrow(data), replace = TRUE)
    boot_data <- data[indices, ]
    
    # fit the models on the bootstrap sample
    boot_mod_X_to_Y <- glm(formula = formula_X_to_Y, data = boot_data, family = binomial())
    boot_mod_X_to_Me_1 <- glm(formula = formula_X_to_Me_1, data = boot_data)
    boot_mod_X_to_Me_2 <- glm(formula = formula_X_to_Me_2, data = boot_data)
    boot_mod_X_to_Me_3 <- glm(formula = formula_X_to_Me_3, data = boot_data)
    boot_mod_XM_to_Y <- glm(formula = formula_XM_to_Y, data = boot_data, family = binomial())
    
    # Extract coefficients
    boot_coef_X_to_Y <- coef(boot_mod_X_to_Y)
    boot_coef_X_to_Me_1 <- coef(boot_mod_X_to_Me_1)
    boot_coef_X_to_Me_2 <- coef(boot_mod_X_to_Me_2)
    boot_coef_X_to_Me_3 <- coef(boot_mod_X_to_Me_3)
    boot_coef_XM_to_Y <- coef(boot_mod_XM_to_Y)
    
    new_df <- data.frame(
                boot = j,
                b1 = boot_coef_X_to_Y[X_name],
                b21 = boot_coef_X_to_Me_1[X_name],
                b22 = boot_coef_X_to_Me_2[X_name],
                b23 = boot_coef_X_to_Me_3[X_name],
                b3 = boot_coef_XM_to_Y[X_name],
                b41 = boot_coef_XM_to_Y[Me_1],
                b42 = boot_coef_XM_to_Y[Me_2],
                b43 = boot_coef_XM_to_Y[Me_3],
                stringsAsFactors = FALSE )
  
    jhs_nRS_multi <- rbind(jhs_nRS_multi, new_df)
}

write.csv(jhs_nRS_multi, "../../results/reordered_results/mediation/multi_mediator/jhs_nRS_multi.csv", row.names = FALSE) 
```


```{r}
jhs_nRS_multi_sum <- jhs_nRS_multi %>%
  mutate(b_21_41 = b21*b41,
         b_22_42 = b22*b42,
         b_23_43 = b23*b43) %>%
  summarise(b1_mean = mean(b1),
           b1_ci_lwr = quantile(b1, 0.025),
           b1_ci_upr = quantile(b1, 0.975),
            
            b21_mean = mean(b21),
            b21_ci_lwr = quantile(b21, 0.025),
            b21_ci_upr = quantile(b21, 0.975),
            
            b22_mean = mean(b22),
            b22_ci_lwr = quantile(b22, 0.025),
            b22_ci_upr = quantile(b22, 0.975),
           
           b23_mean = mean(b23),
           b23_ci_lwr = quantile(b23, 0.025),
            b23_ci_upr = quantile(b23, 0.975),
            
            b3_mean = mean(b3),
            b3_ci_lwr = quantile(b3, 0.025),
            b3_ci_upr = quantile(b3, 0.975),
            
            b41_mean = mean(b41),
            b41_ci_lwr = quantile(b41, 0.025),
            b41_ci_upr = quantile(b41, 0.975),
           
           b42_mean = mean(b42),
            b42_ci_lwr = quantile(b42, 0.025),
            b42_ci_upr = quantile(b42, 0.975),
           
           b43_mean = mean(b43),
            b43_ci_lwr = quantile(b43, 0.025),
            b43_ci_upr = quantile(b43, 0.975),
         
            b_21_41_mean = mean(b_21_41),
            b_21_41_ci_lwr = quantile(b_21_41, 0.025),
            b_21_41_ci_upr = quantile(b_21_41, 0.975),
           
           b_22_42_mean = mean(b_22_42),
            b_22_42_ci_lwr = quantile(b_22_42, 0.025),
            b_22_42_ci_upr = quantile(b_22_42, 0.975),
           
           b_23_43_mean = mean(b_23_43),
            b_23_43_ci_lwr = quantile(b_23_43, 0.025),
            b_23_43_ci_upr = quantile(b_23_43, 0.975)) %>%
  mutate(
    TotalEffect_b1 = sprintf("%.3f [%.3f, %.3f]", b1_mean, b1_ci_lwr, b1_ci_upr),
    b21 = sprintf("%.3f [%.3f, %.3f]", b21_mean, b21_ci_lwr, b21_ci_upr),
    b22 = sprintf("%.3f [%.3f, %.3f]", b22_mean, b22_ci_lwr, b22_ci_upr),
    b23 = sprintf("%.3f [%.3f, %.3f]", b23_mean, b23_ci_lwr, b23_ci_upr),
    DirectEffect_b3 = sprintf("%.3f [%.3f, %.3f]", b3_mean, b3_ci_lwr, b3_ci_upr),
    b41 = sprintf("%.3f [%.3f, %.3f]", b41_mean, b41_ci_lwr, b41_ci_upr),
    b42 = sprintf("%.3f [%.3f, %.3f]", b42_mean, b42_ci_lwr, b42_ci_upr),
    b43 = sprintf("%.3f [%.3f, %.3f]", b43_mean, b43_ci_lwr, b43_ci_upr),
    IndirectEffect_b21_b41 = sprintf("%.3f [%.3f, %.3f]", b_21_41_mean, b_21_41_ci_lwr, b_21_41_ci_upr),
    IndirectEffect_b22_b42 = sprintf("%.3f [%.3f, %.3f]", b_22_42_mean, b_22_42_ci_lwr, b_22_42_ci_upr),
    IndirectEffect_b23_b43 = sprintf("%.3f [%.3f, %.3f]", b_23_43_mean, b_23_43_ci_lwr, b_23_43_ci_upr)
  )

write.csv(jhs_nRS_multi_sum, "../../results/reordered_results/mediation/multi_mediator/jhs_nRS_multi_sum.csv", row.names = FALSE) 
```


#### JHS nSES -> FamIncome, PhysAct, currentSmoker
```{r, warning=FALSE, message=FALSE}
jhs_nSES_multi <- data.frame(
  boot = numeric(),
  b1 = numeric(),
  b21 = numeric(),
  b22 = numeric(),
  b23 = numeric(),
  b3 = numeric(),
  b41 = numeric(),
  b42 = numeric(),
  b43 = numeric(),
  stringsAsFactors = FALSE 
)

# model setting
data <- jhs_std
Y_name <- "cvd_10y_HF"
X_name <- "nSES"
Me_1 <- "FamIncome"
Me_2 <- "PhysAct"
Me_3 <- "currentSmoker"
Z <- c("age","gender","Diabetes", "hdl","totchol","sbp")
  
# generate formula for mediation test
formula_X_to_Y <- paste(Y_name, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Y <- as.formula(formula_X_to_Y)

formula_X_to_Me_1 <- paste(Me_1, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_1 <- as.formula(formula_X_to_Me_1)

formula_X_to_Me_2 <- paste(Me_2, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_2 <- as.formula(formula_X_to_Me_2)

formula_X_to_Me_3 <- paste(Me_3, "~", X_name, "+", paste(Z, collapse=" + "))
formula_X_to_Me_3 <- as.formula(formula_X_to_Me_3)

formula_XM_to_Y <- paste(Y_name, "~", X_name, "+", Me_1, "+", Me_2, "+", Me_3, "+", paste(Z, collapse=" + "))
formula_XM_to_Y <- as.formula(formula_XM_to_Y)

# Bootstrap to get CIs
n_boot <- 1000
set.seed(123)

for (j in 1:n_boot) {
    # Resample 
    indices <- sample(1:nrow(data), replace = TRUE)
    boot_data <- data[indices, ]
    
    # fit the models on the bootstrap sample
    boot_mod_X_to_Y <- glm(formula = formula_X_to_Y, data = boot_data, family = binomial())
    boot_mod_X_to_Me_1 <- glm(formula = formula_X_to_Me_1, data = boot_data)
    boot_mod_X_to_Me_2 <- glm(formula = formula_X_to_Me_2, data = boot_data)
    boot_mod_X_to_Me_3 <- glm(formula = formula_X_to_Me_3, data = boot_data)
    boot_mod_XM_to_Y <- glm(formula = formula_XM_to_Y, data = boot_data, family = binomial())
    
    # Extract coefficients
    boot_coef_X_to_Y <- coef(boot_mod_X_to_Y)
    boot_coef_X_to_Me_1 <- coef(boot_mod_X_to_Me_1)
    boot_coef_X_to_Me_2 <- coef(boot_mod_X_to_Me_2)
    boot_coef_X_to_Me_3 <- coef(boot_mod_X_to_Me_3)
    boot_coef_XM_to_Y <- coef(boot_mod_XM_to_Y)
    
    new_df <- data.frame(
                boot = j,
                b1 = boot_coef_X_to_Y[X_name],
                b21 = boot_coef_X_to_Me_1[X_name],
                b22 = boot_coef_X_to_Me_2[X_name],
                b23 = boot_coef_X_to_Me_3[X_name],
                b3 = boot_coef_XM_to_Y[X_name],
                b41 = boot_coef_XM_to_Y[Me_1],
                b42 = boot_coef_XM_to_Y[Me_2],
                b43 = boot_coef_XM_to_Y[Me_3],
                stringsAsFactors = FALSE )
  
    jhs_nSES_multi <- rbind(jhs_nSES_multi, new_df)
}

write.csv(jhs_nSES_multi, "../../results/reordered_results/mediation/multi_mediator/jhs_nSES_multi.csv", row.names = FALSE) 
```


```{r}
jhs_nSES_multi_sum <- jhs_nSES_multi %>%
  mutate(b_21_41 = b21*b41,
         b_22_42 = b22*b42,
         b_23_43 = b23*b43) %>%
  summarise(b1_mean = mean(b1),
           b1_ci_lwr = quantile(b1, 0.025),
           b1_ci_upr = quantile(b1, 0.975),
            
            b21_mean = mean(b21),
            b21_ci_lwr = quantile(b21, 0.025),
            b21_ci_upr = quantile(b21, 0.975),
            
            b22_mean = mean(b22),
            b22_ci_lwr = quantile(b22, 0.025),
            b22_ci_upr = quantile(b22, 0.975),
           
           b23_mean = mean(b23),
           b23_ci_lwr = quantile(b23, 0.025),
            b23_ci_upr = quantile(b23, 0.975),
            
            b3_mean = mean(b3),
            b3_ci_lwr = quantile(b3, 0.025),
            b3_ci_upr = quantile(b3, 0.975),
            
            b41_mean = mean(b41),
            b41_ci_lwr = quantile(b41, 0.025),
            b41_ci_upr = quantile(b41, 0.975),
           
           b42_mean = mean(b42),
            b42_ci_lwr = quantile(b42, 0.025),
            b42_ci_upr = quantile(b42, 0.975),
           
           b43_mean = mean(b43),
            b43_ci_lwr = quantile(b43, 0.025),
            b43_ci_upr = quantile(b43, 0.975),
         
            b_21_41_mean = mean(b_21_41),
            b_21_41_ci_lwr = quantile(b_21_41, 0.025),
            b_21_41_ci_upr = quantile(b_21_41, 0.975),
           
           b_22_42_mean = mean(b_22_42),
            b_22_42_ci_lwr = quantile(b_22_42, 0.025),
            b_22_42_ci_upr = quantile(b_22_42, 0.975),
           
           b_23_43_mean = mean(b_23_43),
            b_23_43_ci_lwr = quantile(b_23_43, 0.025),
            b_23_43_ci_upr = quantile(b_23_43, 0.975)) %>%
  mutate(
    TotalEffect_b1 = sprintf("%.3f [%.3f, %.3f]", b1_mean, b1_ci_lwr, b1_ci_upr),
    b21 = sprintf("%.3f [%.3f, %.3f]", b21_mean, b21_ci_lwr, b21_ci_upr),
    b22 = sprintf("%.3f [%.3f, %.3f]", b22_mean, b22_ci_lwr, b22_ci_upr),
    b23 = sprintf("%.3f [%.3f, %.3f]", b23_mean, b23_ci_lwr, b23_ci_upr),
    DirectEffect_b3 = sprintf("%.3f [%.3f, %.3f]", b3_mean, b3_ci_lwr, b3_ci_upr),
    b41 = sprintf("%.3f [%.3f, %.3f]", b41_mean, b41_ci_lwr, b41_ci_upr),
    b42 = sprintf("%.3f [%.3f, %.3f]", b42_mean, b42_ci_lwr, b42_ci_upr),
    b43 = sprintf("%.3f [%.3f, %.3f]", b43_mean, b43_ci_lwr, b43_ci_upr),
    IndirectEffect_b21_b41 = sprintf("%.3f [%.3f, %.3f]", b_21_41_mean, b_21_41_ci_lwr, b_21_41_ci_upr),
    IndirectEffect_b22_b42 = sprintf("%.3f [%.3f, %.3f]", b_22_42_mean, b_22_42_ci_lwr, b_22_42_ci_upr),
    IndirectEffect_b23_b43 = sprintf("%.3f [%.3f, %.3f]", b_23_43_mean, b_23_43_ci_lwr, b_23_43_ci_upr)
  )

write.csv(jhs_nSES_multi_sum, "../../results/reordered_results/mediation/multi_mediator/jhs_nSES_multi_sum.csv", row.names = FALSE) 

```





### plotting


### sobel test
```{r}
sobel_df <- data.frame(
  Data = character(),
  Y = character(),
  X = character(),
  mediator = character(),
  b1 = numeric(),
  b1_pvalue = numeric(),
  b2 = numeric(),
  b2_pvalue = numeric(),
  b3 = numeric(),
  b3_pvalue = numeric(),
  b4 = numeric(),
  b4_pvalue = numeric(),
  b2_b4 = numeric(),
  sobel_pvalue = numeric()
)

for (combo in all_combos) {
  
  # get values
  data_name = combo$data
  data <- named_datasets[[data_name]]
  Y <- combo$Y
  X <- combo$X
  M <- combo$M
  Z <- combo$covariates_final
  
  # generate formula
  ## formula for generating weights
  formula_ps_no_str <- paste(X, "~ 1")
  formula_ps_no <- as.formula(formula_ps_no_str)
  formula_ps_str <- paste(X, "~ ", M, "+" ,paste(Z, collapse=" + "))
  formula_ps <- as.formula(formula_ps_str)
  ## formula for mediation test
  formula_X_to_Y <- paste(Y, "~",X, "+", paste(Z, collapse=" + "))
  formula_X_to_Y <- as.formula(formula_X_to_Y)
  formula_X_to_M <- paste(M, "~", X)
  formula_X_to_M <- as.formula(formula_X_to_M)
  formula_XM_to_Y <- paste(Y, "~",X, "+", M, "+", paste(Z, collapse=" + "))
  formula_XM_to_Y <- as.formula(formula_XM_to_Y)
  
  # estimate weights
  denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
  numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
  propensity_density = numerator / denominator
  threshold <- quantile(propensity_density, 0.99)
  data <- data[propensity_density <= threshold, ]
  propensity_density <- propensity_density[propensity_density <= threshold]
  
  # run the 3 models
  
  ## Y ~ X + Z
  mod_X_to_Y <- glm(formula = formula_X_to_Y, data = data, family = binomial(), weights = propensity_density)
  summary_X_to_Y <- summary(mod_X_to_Y)
  coef_X_to_Y <- summary_X_to_Y$coefficients
  ## M ~ X
  mod_X_to_M <- glm(formula = formula_X_to_M, data = data, weights = propensity_density)
  summary_X_to_M <- summary(mod_X_to_M)
  coef_X_to_M <- summary_X_to_M$coefficients
  se_X_to_M <- summary_X_to_M$coefficients[X, "Std. Error"]
  ## Y ~ X + M +Z
  mod_XM_to_Y <- glm(formula = formula_XM_to_Y, data = data, family = binomial(), weights = propensity_density)
  summary_XM_to_Y <- summary(mod_XM_to_Y)
  coef_XM_to_Y <- summary_XM_to_Y$coefficients
  se_XM_to_Y <- summary_XM_to_Y$coefficients[M, "Std. Error"]

  # extract and calculate model result
  b1 <- round(coef_X_to_Y[X, "Estimate"], digits = 5)
  b1_pvalue <- round(coef_X_to_Y[X, "Pr(>|z|)"], digits = 5)
  b2 <- round(coef_X_to_M[X, "Estimate"], digits = 5)
  b2_pvalue <- round(coef_X_to_M[X, "Pr(>|t|)"], digits = 5)
  b3 <- round(coef_XM_to_Y[X, "Estimate"], digits = 5)
  b3_pvalue <- round(coef_XM_to_Y[X, "Pr(>|z|)"], digits = 5)
  b4 <- round(coef_XM_to_Y[M, "Estimate"], digits = 5)
  b4_pvalue <- round(coef_XM_to_Y[M, "Pr(>|z|)"], digits = 5)
  b2_b4 <- round(b2 * b4, digits = 5)
  
  sobel_se <- sqrt((b4^2 * se_X_to_M^2) + (b2^2 * se_XM_to_Y^2))
  sobel_z <- (b2 * b4) / sobel_se
  sobel_pvalue <- 2 * (1 - pnorm(abs(sobel_z)))
  
  # store result
  new_df <- data.frame(
    Data = data_name,
    Y = Y,
    X = X,
    mediator = M,
    b1 = b1,
    b1_pvalue = b1_pvalue,
    b2 = b2,
    b2_pvalue = b2_pvalue,
    b3 = b3,
    b3_pvalue = b3_pvalue,
    b4 = b4,
    b4_pvalue = b4_pvalue,
    b2_b4 = b2_b4,
    sobel_pvalue = sobel_pvalue
  )
  
  sobel_df <- rbind(sobel_df, new_df)

}


# write.csv(sobel_df, "../results/mediation_result/mediation_sobel_240816.csv", row.names = FALSE)
```


significant mediators 
```{r}
coef_ci <- read.csv("../results/mediation_result/mediation_bootstrap_ci_240816.csv")
sobel_df <- read.csv("../results/mediation_result/mediation_sobel_240816.csv")

# Baron and Kenny
# Y ~ X (b1); M ~ X (b2); Y ~ M (b4)
med_BK_ci <- coef_ci %>%
  filter((b1_ci_lwr * b1_ci_upr > 0) & (b2_ci_lwr * b2_ci_upr > 0) & (b4_ci_lwr * b4_ci_upr > 0))
## 0 result
med_BK_coef_sobel <- sobel_df %>%
  filter((b1_pvalue <= 0.05) & (b2_pvalue <= 0.05) & (b4_pvalue <= 0.05))
## 2 results
med_BK_sobel <- sobel_df %>% 
  filter(sobel_pvalue <= 0.05)
## 6 rsults

# Zhao
# b2*b4
med_zhao <- coef_ci %>%
  filter(b2_b4_ci_lwr * b2_b4_ci_upr > 0)
## 3 results

# MacKinnon
# M ~ X (b2); Y ~ M (b4)
med_mackinnon <- coef_ci %>%
  filter((b2_ci_lwr * b2_ci_upr > 0) & (b4_ci_lwr * b4_ci_upr > 0))
## 3 results
```




### alternative mediation methods

#### mediation package

tests
```{r}
library(mediation)

# features
#all_combos[[41]]
```

```{r}
# M ~ X + Z
med.fit <- glm(FamIncome ~ nRS + age + gender + Diabetes + hdl + totchol + sbp, data = jhs_std) 
# Y ~ X + M + Z
out.fit <- glm(cvd_10y_HF ~ FamIncome + nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std, family = binomial("probit")) 

# estimation comparing nRS = 0/1 
# med01.out <- mediate(med.fit, out.fit, treat = "nRS", mediator = "FamIncome", 
#                      control.value = 0, treat.value = 1, boot = TRUE, sims = 100) 
med.out <- mediate(med.fit, out.fit, treat = "nRS", mediator = "FamIncome", boot = TRUE, sims = 100) 

# result
summary(med.out)
plot(med.out)
```

#### standardize the ordinal categorical mediators

jhs nSES FamIncome
```{r}
# Y ~ X + Z
TE.fit <- glm(cvd_10y_HF ~ nSES + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std, family = binomial()) 
summary(TE.fit)

# M ~ X + Z
med.fit <- glm(FamIncome ~ nSES + age + gender + Diabetes + hdl + totchol + sbp, data = jhs_std) 
summary(med.fit)

# Y ~ X + M + Z
out.fit <- glm(cvd_10y_HF ~ FamIncome + nSES + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std, family = binomial()) 
summary(out.fit)

cat("Total Effect =", coef(TE.fit)["nSES"], "\n",
    "Direct Effect =", coef(out.fit)["nSES"], "\n",
    "Indirect Effect =", coef(med.fit)["nSES"] * coef(out.fit)["FamIncome"], "\n")
```


MESA nSES currenSmoker as binary
```{r}
mesa_test <- mesa_std %>%
  mutate(currentSmoker = as.factor(currentSmoker - 1),
         alc = factor(alc - 1))

# Y ~ X + Z
TE.fit <- glm(cvd_10y_HF ~ nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_test, family = binomial()) 
summary(TE.fit)

# M ~ X + Z
med.fit <- glm(currentSmoker ~ nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_test, family = binomial()) 
summary(med.fit)

# Y ~ X + M + Z
out.fit <- glm(cvd_10y_HF ~ currentSmoker + nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_test, family = binomial()) 
summary(out.fit)

cat("Total Effect =", coef(TE.fit)["nSES"], "\n",
    "Direct Effect =", coef(out.fit)["nSES"], "\n",
    "Indirect Effect =", coef(med.fit)["nSES"] * coef(out.fit)[2], "\n")
```

mesa nSES currentSmoker as continuous
```{r}

# Y ~ X + Z
TE.fit <- glm(cvd_10y_HF ~ nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_std, family = binomial()) 
summary(TE.fit)

# M ~ X + Z
med.fit <- glm(currentSmoker ~ nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_std) 
summary(med.fit)

# Y ~ X + M + Z
out.fit <- glm(cvd_10y_HF ~ currentSmoker + nSES + age + gender + Diabetes + hdl + totchol + sbp + race + site, 
               data = mesa_std, family = binomial()) 
summary(out.fit)

cat("Total Effect =", coef(TE.fit)["nSES"], "\n",
    "Direct Effect =", coef(out.fit)["nSES"], "\n",
    "Indirect Effect =", coef(med.fit)["nSES"] * coef(out.fit)[2], "\n")
```


multiple mediators
jhs nRS -> FamIncome, PhysAct, currentSmoker
```{r}
# Y ~ X + Z
TE.fit <- glm(cvd_10y_HF ~ nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std, family = binomial()) 
summary(TE.fit)

# M ~ X + Z
med.fit1 <- glm(FamIncome ~ nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std) 
summary(med.fit1)
med.fit2 <- glm(PhysAct ~ nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std) 
summary(med.fit2)
med.fit3 <- glm(currentSmoker ~ nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std) 
summary(med.fit3)

# Y ~ X + M + Z
out.fit <- glm(cvd_10y_HF ~ FamIncome + PhysAct + currentSmoker + nRS + age + gender + Diabetes + hdl + totchol + sbp, 
               data = jhs_std, family = binomial()) 
summary(out.fit)

# cat("Total Effect =", coef(TE.fit)["nRS"], "\n",
#     "Direct Effect =", coef(out.fit)["nRS"], "\n",
#     "Indirect Effect =", coef(med.fit)["nRS"] * coef(out.fit)[2], "\n")
```




