---
title: "Generalized IPW"
author: "Xiaoting Chen"
date: "2024-05-24"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is translated from the Generalized_IPW.ipynb file and expanded on the spline function and prediction curve generation.\


## Prep

### packages
```{r, warning=FALSE, message=FALSE}
library(readr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(stats) # handle normal distribution calculations
library(splines) # using splines
library(mgcv)
library(DescTools) # calculate mode
library(tableone)
library(cobalt)
```

### data

```{r}
all_combos <- readRDS("all_combos.rds")

# read in moderation and mediation results
significant_moderation <- read.csv("../results/moderation_result/moderation_result.csv") %>%
  filter(significant == "Y") %>%
  select(1:4) %>%
  rename(data_name = Data,
         M = moderator)

significant_mediation <- read.csv("../results/mediation_result/significant_mediat.csv")
```

```{r}
analysis_feature <- c('cvd_10y_HF', 'cvd_10y_noHF', 'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp',
                      'site', 'race')

ind_feature_map <- c('0' = 1, '1' = 2, '2' = 3)
```

```{r}
mesa_std <- read.csv('../data_processed/MESA/mesa_std.csv')

mesa_std <- mesa_std %>%
  select(analysis_feature) %>%
  mutate(# # treat ind behavirol variables as conitnuous, avoid partial significant 
    Diabetes = as.factor(Diabetes),
         site = as.factor(site),
         race = as.factor(race)) %>%
  mutate( # recode to avoid 0
    nutrition = recode(nutrition, !!!ind_feature_map),
    PhysAct = recode(PhysAct, !!!ind_feature_map),
    currentSmoker = recode(currentSmoker, !!!ind_feature_map),
    alc = recode(alc, !!!ind_feature_map)
  )

mesa_bla_std <- mesa_std[mesa_std$race == 1, ] %>%
  select(-race)
```

```{r}
jhs_std <- read.csv('../data_processed/JHS/jhs_std.csv')

analysis_feature <- c('cvd_10y_HF', 'cvd_10y_noHF', 'nSES', 'nFavFood', 'nPhysFac', 'nRS',
                      'FamIncome', 'nutrition', 'PhysAct', 'currentSmoker', 'alc',
                      'age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp')

jhs_std <- jhs_std %>%
  select(analysis_feature) %>%
  mutate(# # treat ind behavirol variables as conitnuous, avoid partial significant 
    Diabetes = as.factor(Diabetes)) %>%
  mutate( # recode to avoid 0
    nutrition = recode(nutrition, !!!ind_feature_map),
    PhysAct = recode(PhysAct, !!!ind_feature_map),
    currentSmoker = recode(currentSmoker, !!!ind_feature_map),
    alc = recode(alc, !!!ind_feature_map)
  )
```

```{r}
named_datasets <- list(mesa = mesa_std, mesa_bla = mesa_bla_std,jhs = jhs_std)
names(named_datasets) <- c("mesa", "mesa_bla","jhs")
```

### function    
```{r}
conditional_densities <- function(data, treatment, formula_ps_no_con, formula_ps_con, use_confounders = TRUE) {

  formula <- if(use_confounders) formula_ps_con else formula_ps_no_con
  
  # Fit the linear model
  model <- lm(as.formula(formula), data = data)
  
  # Calculate the fitted values and standard deviation of residuals
  fitted_values <- fitted(model)
  resid_std <- sd(resid(model))
  
  # Calculate the density of treatment under a normal distribution with parameters from the model
  densities <- dnorm(data[[treatment]], mean = fitted_values, sd = resid_std)
  
  # Return the densities as a vector indexed similarly to the fitted values
  return(setNames(densities, names(fitted_values)))
}

```


## loop to generate result

### estimate
```{r, warning=FALSE}
report_df <- data.frame(
  Data = character(),
  Y = numeric(),
  X = numeric(),
  intercept = numeric(),
  intercept_pvalue = numeric(),
  X_coef = numeric(),
  X_pvalue = numeric(),
  X_ci_lwr = numeric(),
  X_ci_upr = numeric(),
  moderator = numeric(),
  moderator_coef = numeric(),
  moderator_pvalue = numeric(),
  moderator_ci_lwr = numeric(),
  moderator_ci_upr = numeric(),
  interaction_coef = numeric(),
  interaction_pvalue = numeric(),
  interact_ci_lwr = numeric(),
  interact_ci_upr = numeric(),
  stringsAsFactors = FALSE 
)


for (combo in all_combos) {
  
  # get values
  data_name = combo$data
  data <- named_datasets[[data_name]]
  
  Y <- combo$Y
  X <- combo$X
  M <- combo$M
  Z <- combo$covariates_final
  
  # generate formula; add interaction based on moderation result
  matches <- which(significant_moderation$data_name == data_name & 
                     significant_moderation$Y == Y & 
                     significant_moderation$X == X & 
                     significant_moderation$M == M)

  if(length(matches) > 0) { # exist moderation, add interaction term in outcome model
    
    formula_ps_no_str <- paste(X, "~ 1")
    formula_ps_no <- as.formula(formula_ps_no_str)
    
    formula_ps_str <- paste(X, "~ ", M, "+" ,paste(Z, collapse=" + "))
    formula_ps <- as.formula(formula_ps_str)
    
    formula_outcome_str <- paste(Y, "~",X,"+",M,"+",paste(X, M, sep=":"))
    formula_outcome <- as.formula(formula_outcome_str)
    
    moderator <- M
    
  } else { # no moderation
    
    formula_ps_no_str <- paste(X, "~ 1")
    formula_ps_no <- as.formula(formula_ps_no_str)
    
    formula_ps_str <- paste(X, "~ ", M, "+" ,paste(Z, collapse=" + "))
    formula_ps <- as.formula(formula_ps_str)
    
    formula_outcome_str <- paste(Y, "~",X)
    formula_outcome <- as.formula(formula_outcome_str)
    
    moderator <- NA
  }
  
  
  # estimate
  
  denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
  numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
  propensity_density = numerator / denominator
  threshold <- quantile(propensity_density, 0.99)
  data <- data[propensity_density <= threshold, ]
  propensity_density <- propensity_density[propensity_density <= threshold]
  outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
  summary_model <- summary(outcome_mod)

  # store model result
  # there will be duplicate rows of results since we dont consider M unless significant moderation
  # remove duplicates later

  coefficients <- summary_model$coefficients
  ci <- confint(outcome_mod)
  
  intercept <- round(coefficients["(Intercept)", "Estimate"], digits = 5)  ## common parts
  intercept_pvalue <- round(coefficients["(Intercept)", "Pr(>|z|)"], digits = 5)
  
  X_coef <- round(coefficients[X, "Estimate"], digits = 5)
  X_pvalue <- round(coefficients[X, "Pr(>|z|)"], digits = 5)
  X_ci_lwr <- round(ci[X,"2.5 %"],digits = 5)
  X_ci_upr <- round(ci[X,"97.5 %"],digits = 5)

  if(is.na(moderator)) {
    
    moderator_coef <- NA
    moderator_pvalue <- NA
    moderator_ci_lwr <- NA
    moderator_ci_upr <- NA
    
    interaction_coef <- NA
    interaction_pvalue <- NA
    interact_ci_lwr <- NA
    interact_ci_upr <- NA
    
  } else{
    
    moderator_coef <- round(coefficients[moderator, "Estimate"], digits = 5)
    moderator_pvalue <- round(coefficients[moderator, "Pr(>|z|)"], digits = 5)
    moderator_ci_lwr <- round(ci[moderator,"2.5 %"],digits = 5)
    moderator_ci_upr <- round(ci[moderator,"97.5 %"],digits = 5)
      
    interact_term_name <- tail(rownames(coefficients), 1)
    interaction_coef <- round(coefficients[interact_term_name, "Estimate"], digits = 5)
    interaction_pvalue <- round(coefficients[interact_term_name, "Pr(>|z|)"], digits = 5)
    interact_ci_lwr <- round(ci[interact_term_name,"2.5 %"],digits = 5)
    interact_ci_upr <- round(ci[interact_term_name,"97.5 %"],digits = 5)
  }
  
  new_df <- data.frame(
              Data = data_name,
              Y = Y,
              X = X,
              intercept = intercept,
              intercept_pvalue = intercept_pvalue,
              X_coef = X_coef,
              X_pvalue = X_pvalue,
              X_ci_lwr = X_ci_lwr,
              X_ci_upr = X_ci_upr,
              moderator = moderator,
              moderator_coef = moderator_coef,
              moderator_pvalue = moderator_pvalue,
              moderator_ci_lwr = moderator_ci_lwr,
              moderator_ci_upr = moderator_ci_upr,
              interaction_coef = interaction_coef,
              interaction_pvalue = interaction_pvalue,
              interact_ci_lwr = interact_ci_lwr,
              interact_ci_upr = interact_ci_upr,
              stringsAsFactors = FALSE )

  report_df <- rbind(report_df, new_df)

}

```

```{r}
# exclude duplicate rows
ipw_moderate <- report_df %>%
  distinct() 

ipw_moderate <- ipw_moderate[-c(2,8,13,18,21,24,27), ]# exclude the rows that should have been moderated
```

### modify the mediated effect
```{r}
ipw_mediate <- ipw_moderate 
ipw_mediate$mediator <- NA
ipw_mediate$mediator_coef <- NA
ipw_mediate$mediator_pvalue <- NA
ipw_mediate$mediator_ci_lwr <- NA
ipw_mediate$mediator_ci_upr <- NA

# mesa nSES being mediated
ipw_mediate[1,"mediator"] <- "nutrition&PhysAct"  
ipw_mediate[5,"mediator"] <- "nutrition"
```

```{r}
data <- mesa_std
treatment = "nSES"
formula_ps_no = "nSES ~ 1"

formula_ps = "nSES ~ 1 + nRS + nFavFood + nPhysFac + \
                FamIncome + nutrition + PhysAct + currentSmoker + alc + \
                sbp + Diabetes + hdl + totchol + age + gender +\
                site + race"

formula_outcome_logit = 'cvd_10y_HF ~ 1 + nSES + nutrition + PhysAct'
```

Step 1: estimate propensity density
```{r}
denominator = conditional_densities(data, treatment, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,treatment, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator

# exclude extreme values
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]
propensity_density <- propensity_density[propensity_density <= threshold]
```

Step 2: Outcome model
```{r, warning=FALSE}
modelA <- glm(formula = formula_outcome_logit, data = data, family = binomial(), weights = propensity_density)
summary(modelA)
confint(modelA)
```

```{r}
ipw_mediate[1,"intercept"] <- -1.29558
ipw_mediate[1,"X_coef"] <- 0.12714
ipw_mediate[1,"X_pvalue"] <- 0.04996
ipw_mediate[1,"X_ci_lwr"] <- 0.001158415
ipw_mediate[1,"X_ci_upr"] <- 0.2555166
ipw_mediate[1,"mediator_coef"] <- "-0.34408; -0.15921"
ipw_mediate[1,"mediator_pvalue"] <- "0.00319; 0.01471"
ipw_mediate[1,"mediator_ci_lwr"] <- "-0.575831300; -0.286160205"
ipw_mediate[1,"mediator_ci_upr"] <- "-0.1180496; -0.0301479"

ipw_mediate[5,"intercept"] <- -1.7693
ipw_mediate[5,"X_coef"] <- 0.2386
ipw_mediate[5,"X_pvalue"] <- 0.000937
ipw_mediate[5,"X_ci_lwr"] <- 0.09896727
ipw_mediate[5,"X_ci_upr"] <- 0.3817504
ipw_mediate[5,"mediator_coef"] <- -0.4484
ipw_mediate[5,"mediator_pvalue"] <- 0.000552
ipw_mediate[5,"mediator_ci_lwr"] <- -0.70724600
ipw_mediate[5,"mediator_ci_upr"] <- -0.1977654
```


```{r}
write.csv(ipw_mediate, "../results/generalized_IPW/ipw_final_0530.csv", row.names = FALSE)
```


```{r}
ipw_significant <- read.csv("/Users/cxt/Documents/research/CVD/Effect-of-nSES-on-CVD/results/generalized_IPW/ipw_final_0530.csv") %>%
  filter(X_pvalue <= 0.05) %>%
  filter(Data != "mesa_bla") 

write.csv(ipw_significant, "../results/generalized_IPW/ipw_significant_0530.csv", row.names = FALSE)
```


### plot prediction curve for significant results
```{r}
ipw_final <- read.csv("../results/generalized_IPW/ipw_final_0528.csv")

plot_ipw <- ipw_final %>%
  filter(X_pvalue <= 0.05) 
```

```{r}
for (i in 1:nrow(plot_ipw)) {
  print(i)
  
  # extract values
  data_name <- plot_ipw[i, "Data"]
  data <- named_datasets[[data_name]]
  Y <- plot_ipw[i, "Y"]
  X <- plot_ipw[i, "X"]
  moderator <- plot_ipw[i, "moderator"]
  X_coef <- plot_ipw[i, "X_coef"]
  X_pvalue <- plot_ipw[i, "X_pvalue"]
  interaction_coef <- plot_ipw[i, 'interaction_coef']
  interaction_pvalue <- plot_ipw[i, "interaction_pvalue"]
  
  # generate formula 
  
  ## covariates
  base_X <- c("nSES", "nPhysFac", "nFavFood", "nRS")
  base_Z <- c('age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp')
  base_M <- c("FamIncome", "nutrition", "PhysAct", "currentSmoker", "alc")
  
  if (data_name == "mesa") {
    final_Z = c(base_Z, 'race', 'site')
  } else if (data_name == "mesa_bla") {
    final_Z = c(base_Z, 'site')
  } else if (data_name == "jhs") {
    final_Z = base_Z
  }
  
  ## formula
  formula_ps_no <- as.formula(paste(X, "~ 1"))
  
  formula_ps <- as.formula(paste(X, "~ ", paste(setdiff(base_X,X), collapse = "+"),"+",
                                 paste(base_M, collapse=" + "), "+" ,paste(final_Z, collapse=" + ")))
  
  formula_outcome_str <- ifelse(is.na(moderator), 
                            paste(Y, "~",X), 
                            paste(Y, "~",X,"+",moderator,"+",paste(X, moderator, sep=":"))
                            )
  formula_outcome <- as.formula(formula_outcome_str)
  
  # fit model
  denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
  numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
  propensity_density = numerator / denominator
  threshold <- quantile(propensity_density, 0.99)
  data <- data[propensity_density <= threshold, ]    
  propensity_density <- propensity_density[propensity_density <= threshold]
  outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
    
  # Predicted probabilities and plotting
  if (is.na(moderator)) {
    pred_data <- data.frame(seq(from = min(data[[X]]), to = max(data[[X]]), by = 0.1))
    names(pred_data) <- X
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se

    plt <- ggplot(pred_data, aes(x = !!sym(X))) +
      geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.2) +
      geom_line(aes(y = fit), color = "blue") +
      geom_hline(yintercept = 0, color = "black") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()

  } else if (moderator == "alc") {
    X_seq <- seq(from = min(data[[X]]), to = max(data[[X]]), by = 0.1)
    X_length <- length(X_seq)
    pred_data <- data.frame(X = rep(X_seq, 2),
                            alc = c(rep(1, X_length), rep(2, X_length)))
    names(pred_data) <- c(X, "alc")
    
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(alc, levels = c(1,2), labels = c("No", "Yes")), 
                                 color = factor(alc))) +
      geom_line() +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("blue", "red"),  labels = c("No", "Yes"),
                         name = "Alcohol Drinking") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()

  } else if (moderator == "FamIncome") {
    X_seq <- seq(from = min(data[[X]]), to = max(data[[X]]), by = 0.1)
    X_length <- length(X_seq)
    pred_data <- data.frame(X = rep(X_seq, 4),
                            FamIncome = c(rep(1, X_length), rep(2, X_length), rep(3, X_length), rep(4, X_length)))
    names(pred_data) <- c(X, "FamIncome")
    
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(FamIncome, levels = c(1,2,3,4)), 
                                 color = factor(FamIncome))) +
      geom_line() +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("red", "blue", "green", "orange"),  
                         labels = c("$0-11,999", "$12,000-24,999", "$25,000-74,999", "$75,000+"),
                         name = "Family Income") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()

  } else if (moderator == "PhysAct") {
    X_seq <- seq(from = min(data[[X]]), to = max(data[[X]]), by = 0.1)
    X_length <- length(X_seq)
    pred_data <- data.frame(X = rep(X_seq, 3),
                            PhysAct = c(rep(1, X_length), rep(2, X_length), rep(3, X_length)))
    names(pred_data) <- c(X, "PhysAct")
    
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(PhysAct, levels = c(1,2,3)),  
                                 color = factor(PhysAct))) +
      geom_line() +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("blue", "red","green"),  labels = c("Poor", "Intermediate","Ideal"),
                         name = "Physical Activity") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()
  } else {
    print("error")
  }
  
  # export plot
  # ggsave(paste0(data_name,"_",Y,"_",X,".png"), plot = plt,
  #          path = "../results/generalized_IPW/significant_prediction_curve",
  #          width = 8, height = 6, dpi = 300)
}

```


#### plot: X range +- 2sd
```{r}
for (i in 1:nrow(plot_ipw)) {
  print(i)
  
  # extract values
  data_name <- plot_ipw[i, "Data"]
  data <- named_datasets[[data_name]]
  Y <- plot_ipw[i, "Y"]
  X <- plot_ipw[i, "X"]
  moderator <- plot_ipw[i, "moderator"]
  X_coef <- plot_ipw[i, "X_coef"]
  X_pvalue <- plot_ipw[i, "X_pvalue"]
  interaction_coef <- plot_ipw[i, 'interaction_coef']
  interaction_pvalue <- plot_ipw[i, "interaction_pvalue"]
  
  # generate formula 
  
  ## covariates
  base_X <- c("nSES", "nPhysFac", "nFavFood", "nRS")
  base_Z <- c('age', 'gender', 'Diabetes', 'hdl', 'totchol', 'sbp')
  base_M <- c("FamIncome", "nutrition", "PhysAct", "currentSmoker", "alc")
  
  if (data_name == "mesa") {
    final_Z = c(base_Z, 'race', 'site')
  } else if (data_name == "mesa_bla") {
    final_Z = c(base_Z, 'site')
  } else if (data_name == "jhs") {
    final_Z = base_Z
  }
  
  ## formula
  formula_ps_no <- as.formula(paste(X, "~ 1"))
  
  formula_ps <- as.formula(paste(X, "~ ", paste(setdiff(base_X,X), collapse = "+"),"+",
                                 paste(base_M, collapse=" + "), "+" ,paste(final_Z, collapse=" + ")))
  
  formula_outcome_str <- ifelse(is.na(moderator), 
                            paste(Y, "~",X), 
                            paste(Y, "~",X,"+",moderator,"+",paste(X, moderator, sep=":"))
                            )
  formula_outcome <- as.formula(formula_outcome_str)
  
  # fit model
  denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
  numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
  propensity_density = numerator / denominator
  threshold <- quantile(propensity_density, 0.99)
  data <- data[propensity_density <= threshold, ]    
  propensity_density <- propensity_density[propensity_density <= threshold]
  outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
    
  # Predicted probabilities and plotting
  if (is.na(moderator)) {
    pred_data <- data.frame(seq(from = mean(data[[X]])-2*sd(data[[X]]), 
                                to = mean(data[[X]])+2*sd(data[[X]]), by = 0.1))
    names(pred_data) <- X
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se

    plt <- ggplot(pred_data, aes(x = !!sym(X))) +
      geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.2) +
      geom_line(aes(y = fit), color = "blue") +
      geom_hline(yintercept = 0, color = "black") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()

  } else if (moderator == "alc") {
    stats <- data %>%
      group_by(alc) %>%
      summarize(mean_X = mean(!!sym(X), na.rm = TRUE), 
                sd_X = sd(!!sym(X), na.rm = TRUE))
    
    pred_data <- do.call(rbind, lapply(1:2, function(level) {
      mean_X <- stats$mean_X[stats$alc == level]
      sd_X <- stats$sd_X[stats$alc == level]
      X_seq <- seq(from = mean_X - 2 * sd_X, to = mean_X + 2 * sd_X, by = 0.1)
      data.frame(X = X_seq, alc = rep(level, length(X_seq)))
    }))
    names(pred_data) <- c(X, "alc")
    
    preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(alc, levels = c(1,2), labels = c("No", "Yes")), 
                                 color = factor(alc))) +
      geom_line() +
      geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2) +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("blue", "red"),  labels = c("No", "Yes"),
                         name = "Alcohol Drinking") +
      labs(title = "Predicted Probabilities of Positive Outcome",
         x = X, y = "Estimated Probability") +
      theme_minimal()

  } else if (moderator == "FamIncome") {
    stats <- data %>%
      group_by(FamIncome) %>%
      summarize(mean_X = mean(!!sym(X), na.rm = TRUE), 
                sd_X = sd(!!sym(X), na.rm = TRUE))
   
    pred_data <- do.call(rbind, lapply(1:4, function(level) {
      mean_X <- stats$mean_X[stats$FamIncome == level]
      sd_X <- stats$sd_X[stats$FamIncome == level]
      X_seq <- seq(from = mean_X - 2 * sd_X, to = mean_X + 2 * sd_X, by = 0.1)
      data.frame(X = X_seq, FamIncome = rep(level, length(X_seq)))
    }))
    names(pred_data) <- c(X, "FamIncome")
    
    preds <- predict(outcome_mod, pred_data, type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(FamIncome, levels = c(1, 2, 3, 4)),  
                                 color = factor(FamIncome))) +
      geom_line() +
      geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2) +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("red", "blue", "green", "orange"), 
                         labels = c("$0-11,999", "$12,000-24,999", "$25,000-74,999", "$75,000+"),
                         name = "Family Income") +
      labs(title = "Predicted Probabilities of Positive Outcome",
           x = X, y = "Estimated Probability") +
      theme_minimal()


  } else if (moderator == "PhysAct") {
    stats <- data %>%
      group_by(PhysAct) %>%
      summarize(mean_X = mean(!!sym(X), na.rm = TRUE), 
                sd_X = sd(!!sym(X), na.rm = TRUE))
    
    # Generate prediction data
    pred_data <- do.call(rbind, lapply(1:3, function(level) {
      mean_X <- stats$mean_X[stats$PhysAct == level]
      sd_X <- stats$sd_X[stats$PhysAct == level]
      X_seq <- seq(from = mean_X - 2 * sd_X, to = mean_X + 2 * sd_X, by = 0.1)
      data.frame(X = X_seq, PhysAct = rep(level, length(X_seq)))
    }))
    
    # Set column names
    names(pred_data) <- c(X, "PhysAct")
    
    # Get predictions
    preds <- predict(outcome_mod, pred_data, type = "response", se.fit = TRUE)
    se <- preds$se.fit
    pred_data$fit <- preds$fit
    pred_data$lwr <- preds$fit - 1.96 * se
    pred_data$upr <- preds$fit + 1.96 * se
    
    # Plot
    plt <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(PhysAct, levels = c(1, 2, 3)),  
                                 color = factor(PhysAct))) +
      geom_line() +
      geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2) +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("blue", "red", "green"), 
                         labels = c("Poor", "Intermediate", "Ideal"),
                         name = "Physical Activity") +
      labs(title = "Predicted Probabilities of Positive Outcome",
           x = X, y = "Estimated Probability") +
      theme_minimal()
  } else {
    print("error")
  }
  
  # export plot
  ggsave(paste0(data_name,"_",Y,"_",X,".png"), plot = plt,
           path = "../results/generalized_IPW/significant_prediction_curve_0528",
           width = 8, height = 6, dpi = 300)
}

```

```{r}
test_dat <- data %>%
  mutate(PhysAct = factor(PhysAct, levels = c(1, 2, 3)))

# Create the density plot
plt_density <- ggplot(test_dat, aes(x = !!sym("nSES"), fill = PhysAct, color = PhysAct)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("blue", "red", "green"), 
                    labels = c("Poor", "Intermediate", "Ideal"),
                    name = "Physical Activity") +
  scale_color_manual(values = c("blue", "red", "green"), 
                     labels = c("Poor", "Intermediate", "Ideal"),
                     name = "Physical Activity") +
  labs(title = "Density Plot of X by Levels of Physical Activity",
       x = X, y = "Density") +
  theme_minimal()

print(plt_density)
```



## Manually plot (latest)

MESA nSES
```{r, warning=FALSE}
data <- mesa_std
Y <- "cvd_10y_HF"
X <- "nSES"

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ nPhysFac+nFavFood+nRS+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+Diabetes+hdl+totchol+sbp+\
                         race+site")
formula_outcome <- as.formula("cvd_10y_HF ~ nSES + nutrition + PhysAct")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
    
# Predicted probabilities and plotting
## generate pred data
pred_data <- data.frame(seq(from = mean(data[[X]])-2*sd(data[[X]]), 
                                to = mean(data[[X]])+2*sd(data[[X]]), by = 0.1))
names(pred_data) <- X
pred_data$nutrition <- Mode(mesa_std$nutrition)
pred_data$PhysAct <- Mode(mesa_std$PhysAct)

preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
se <- preds$se.fit
pred_data$fit <- preds$fit
pred_data$lwr <- preds$fit - 1.96 * se
pred_data$upr <- preds$fit + 1.96 * se

plt1 <- ggplot(pred_data, aes(x = !!sym(X))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.2) +
  geom_line(aes(y = fit), color = "blue") +
  geom_hline(yintercept = 0, color = "black") +
  labs(x = "Neighborhood Socioeconomic Status Level in MESA", y = "Predicted Probabilities of Experiencing CVD") +
  theme_minimal() +
  theme(
        axis.title.x = element_text(size = 20),  
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 20),  
    axis.text.y = element_text(size = 20)   
      )

plt1

ggsave("mesa_nses.png", plot = plt1,
           path = "../results/generalized_IPW/prediction_curve_final",
           width = 10, height = 8, dpi = 300)

```

JHS nSES
```{r}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nSES"

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ nPhysFac+nFavFood+nRS+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+gender+Diabetes+hdl+totchol+sbp")
formula_outcome <- as.formula("cvd_10y_HF ~ nSES + PhysAct + nSES:PhysAct")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
    
# Predicted probabilities and plotting
## generate pred data
stats <- data %>%
      group_by(PhysAct) %>%
      summarize(mean_X = mean(!!sym(X), na.rm = TRUE), 
                sd_X = sd(!!sym(X), na.rm = TRUE))
    
# Generate prediction data
pred_data <- do.call(rbind, lapply(1:3, function(level) {
  mean_X <- stats$mean_X[stats$PhysAct == level]
  sd_X <- stats$sd_X[stats$PhysAct == level]
  X_seq <- seq(from = mean_X - 2 * sd_X, to = mean_X + 2 * sd_X, by = 0.1)
  data.frame(X = X_seq, PhysAct = rep(level, length(X_seq)))
    }))
    
# Set column names
names(pred_data) <- c(X, "PhysAct")
    
# Get predictions
preds <- predict(outcome_mod, pred_data, type = "response", se.fit = TRUE)
se <- preds$se.fit
pred_data$fit <- preds$fit
pred_data$lwr <- preds$fit - 1.96 * se
pred_data$upr <- preds$fit + 1.96 * se
    
# Plot
plt2 <- ggplot(pred_data, aes(x = !!sym(X), y = fit, 
                                 group = factor(PhysAct, levels = c(1, 2, 3)),  
                                 color = factor(PhysAct))) +
      geom_line() +
      geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2) +
      geom_hline(yintercept = 0, color = "black") +
      scale_color_manual(values = c("blue", "red", "green"), 
                         labels = c("Poor", "Intermediate", "Ideal"),
                         name = "Physical Activity") +
      labs(x = "Neighborhood Socioeconomic Status Level in JHS", y = "Predicted Probabilities of Experiencing CVD") +
  theme_minimal() +
  theme(
        axis.title.x = element_text(size = 20),  
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 20),  
    axis.text.y = element_text(size = 20),
        legend.text = element_text(size = 20),
    legend.title = element_text(size = 20) 
      )
plt2
ggsave("jhs_nses.png", plot = plt2,
           path = "../results/generalized_IPW/prediction_curve_final",
           width = 12, height = 8, dpi = 300)
```

JHS nRS
```{r, warning=FALSE}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nRS"

# generate formula 
formula_ps_no <- as.formula("nRS ~ 1")
formula_ps <- as.formula("nRS ~ nSES+nPhysFac+nFavFood+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+Diabetes+hdl+totchol+sbp")
formula_outcome <- as.formula("cvd_10y_HF ~ nRS")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
outcome_mod <- glm(formula = formula_outcome, data = data, family = binomial(), weights = propensity_density)
    
# Predicted probabilities and plotting
## generate pred data
pred_data <- data.frame(seq(from = mean(data[[X]])-2*sd(data[[X]]), 
                                to = mean(data[[X]])+2*sd(data[[X]]), by = 0.1))
names(pred_data) <- X

preds <- predict(outcome_mod, pred_data,type = "response", se.fit = TRUE)
se <- preds$se.fit
pred_data$fit <- preds$fit
pred_data$lwr <- preds$fit - 1.96 * se
pred_data$upr <- preds$fit + 1.96 * se

plt3 <- ggplot(pred_data, aes(x = !!sym(X))) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.2) +
  geom_line(aes(y = fit), color = "blue") +
  geom_hline(yintercept = 0, color = "black") +
  labs(x = "Neighborhood Racial Segregation Level in JHS", y = "Predicted Probabilities of Experiencing CVD") +
  theme_minimal() +
  theme(
        axis.title.x = element_text(size = 20),  
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 20),  
    axis.text.y = element_text(size = 20)
      )

ggsave("jhs_nrs.png", plot = plt3,
           path = "../results/generalized_IPW/prediction_curve_final",
           width = 10, height = 8, dpi = 300)
```

## Check assumption

dichotomize the neighborhood exposure variables after weight calculation, to examine overlap and balance\\
tutorial from https://ehsanx.github.io/psw/s3.html\\


MESA nSES
```{r, warning=FALSE}
data <- mesa_std
Y <- "cvd_10y_HF"
X <- "nSES"
Z <- c("nPhysFac","nFavFood","nRS",
       "FamIncome","nutrition","PhysAct","currentSmoker","alc",
       "age","Diabetes","hdl","totchol","sbp","gender",
       "race","site")

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ nPhysFac+nFavFood+nRS+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+Diabetes+hdl+totchol+sbp+gender+\
                         race+site")
formula_outcome <- as.formula("cvd_10y_HF ~ nSES + nutrition + PhysAct")
  
# get weights
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density


smd_after <- bal.tab(
  data[, Z], 
  treat = "nSES", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
          threshold = 0.1,  # Preferred threshold for good balance
          var.order = "alphabetical") +
  labs(title = "Pearson correlation between the covariate and treatment before and after Weighting\
  In MESA using n-SES as exposure",
       subtitle = "Dashed line: PC = -0.1 and PC = 0.1 (preferred threshold)")

ggsave("mesa_nses_balance.png", plot = plt,
           path = "../results/generalized_IPW/check_balance",
           width = 8, height = 6, dpi = 300)
```


JHS nSES
```{r}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nSES"
Z <- c("nPhysFac","nFavFood","nRS",
       "FamIncome","nutrition","PhysAct","currentSmoker","alc",
       "age","Diabetes","hdl","totchol","sbp","gender")

# generate formula 
formula_ps_no <- as.formula("nSES ~ 1")
formula_ps <- as.formula("nSES ~ nPhysFac+nFavFood+nRS+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+gender+Diabetes+hdl+totchol+sbp")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density

smd_after <- bal.tab(
  data[, Z], 
  treat = "nSES", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
          threshold = 0.1,  # Preferred threshold for good balance
          var.order = "alphabetical") +
  labs(title = "Pearson correlation between the covariate and treatment before and after Weighting\
  In JHS using n-SES as exposure",
       subtitle = "Dashed line: PC = -0.1 and PC = 0.1 (preferred threshold)")

ggsave("jhs_nses_balance.png", plot = plt,
           path = "../results/generalized_IPW/check_balance",
           width = 8, height = 6, dpi = 300)
```


JHS nRS
```{r}
data <- jhs_std
Y <- "cvd_10y_HF"
X <- "nRS"
Z <- c("nPhysFac","nFavFood","nSES",
       "FamIncome","nutrition","PhysAct","currentSmoker","alc",
       "age","Diabetes","hdl","totchol","sbp","gender")

# generate formula 
formula_ps_no <- as.formula("nRS ~ 1")
formula_ps <- as.formula("nRS ~ nSES+nPhysFac+nFavFood+\
                         FamIncome+nutrition+PhysAct+currentSmoker+alc+\
                         age+gender+Diabetes+hdl+totchol+sbp")
  
# fit model
denominator = conditional_densities(data, X, formula_ps_no,formula_ps, use_confounders=T)
numerator = conditional_densities(data,X, formula_ps_no,formula_ps, use_confounders=F)
propensity_density = numerator / denominator
threshold <- quantile(propensity_density, 0.99)
data <- data[propensity_density <= threshold, ]    
propensity_density <- propensity_density[propensity_density <= threshold]
data$weights <- propensity_density

# using continuous nSES
smd_after <- bal.tab(
  data[, Z], 
  treat = "nRS", 
  data =data,
  weights = "weights",
  un = TRUE
)

# Plotting the SMDs before and after weighting
plt <- love.plot(smd_after, 
          threshold = 0.1,  # Preferred threshold for good balance
          var.order = "alphabetical") +
  labs(title = "Pearson correlation between the covariate and treatment before and after Weighting\
  In JHS using n-Racial Segregation as exposure",
       subtitle = "Dashed line: PC = -0.1 and PC = 0.1 (preferred threshold)")

ggsave("jhs_nrs_balance.png", plot = plt,
           path = "../results/generalized_IPW/check_balance",
           width = 8, height = 6, dpi = 300)
```




